Approximation as Differentiation: Notes

### 2025-05-25 main/higher-order.tex
Onto higher-order skeleton.
Set-wide products. If C has biproducts and set-wide products, then C is CC.

@@

We can alternatively take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
\biproj_Y, \biinj_X, \biinj_Y)$ in any category $\cat{C}$ known to be $\CMon$-enriched, and therefore
providing addition of morphisms; in that case any products or coproducts in $\cat{C}$ are necessarily
biproducts.

>>>>

Although $\Fam(\FinVect)$ nor $\Fam(\LatGal)$ provide an appropriate notion of composition for automatic
differentiation and Galois slicing respectively, neither is Cartesian closed. Both are therefore unsuitable
for interpreting higher-order programs. However the following is a special case of a construction due to
\citet{nunes2023}:

\begin{proposition}
Suppose $\cat{C}$ locally small. If $\cat{C}$ has binary biproducts and all small products, then
$\Fam(\cat{C})$ is Cartesian closed.
\end{proposition}

$\cat{C}$ has all small (set-indexed) products if, for any set $X$ and any $X$-indexed family $A$ of objects
of $\cat{C}$, there exists an object $\prod_{x \in X}A_x$ (also written $\prod_{X} A$) and family of
projection morphisms $\{\eval_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$
and any family of morphisms $\{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product
morphism $\lambda(f)$ making the following diagram commute for every $x \in X$:

\begin{center}
\begin{tikzcd}
   \prod_{X} A \arrow[r, "\eval_x"] & A_x \\
   B \arrow[ru, "f_x"'] \arrow[u, dotted, "\lambda(f)"]
\end{tikzcd}
\end{center}

@@

Alternatively, we can take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
\biproj_Y, \biinj_X, \biinj_Y)$ in any category $\cat{C}$ known to be $\CMon$-enriched, and therefore
providing addition of morphisms; in that case any products or coproducts in $\cat{C}$ are necessarily
biproducts.
